---
title: "R_advanced_P5_Apply"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Import Data

```{r}
Chicago_F = read.csv("./data/Weather Data/Chicago-F.csv", row.names = 1)
NewYork_F = read.csv("./data/Weather Data/NewYork-F.csv", row.names = 1)
Houston_F = read.csv("./data/Weather Data/Houston-F.csv", row.names = 1)
SanFrancisco_F = read.csv("./data/Weather Data/SanFrancisco-F.csv", row.names = 1)
```

```{r}
is.data.frame(Chicago_F)
```
### Let's convert to matrices:

```{r}
Chicago_F = as.matrix(Chicago_F)
NewYork_F = as.matrix(NewYork_F)
Houston_F = as.matrix(Houston_F)
SanFrancisco_F = as.matrix(SanFrancisco_F)
```

check:

```{r}
is.matrix(Chicago_F)
```

## Put all of these data into a list:

```{r}
weather_F = list(Chicago = Chicago_F, Houston = Houston_F, NewYork = NewYork_F, SanFrancisco = SanFrancisco_F) 
```

Try

```{r}
weather_F[[3]] %>% knitr::kable()
weather_F$NewYork %>% knitr::kable()
```

## Apply Family
- `apply()` : use on a matrix, either the rows or columns.
- `tapply()` : use on a vector to extract subgroups and apply a function to them by - use on data frames. Same concept as in group by in SQL.
- `eapply()` : use on an environment.
- `lapply()` : apply a function to elements of a list.
- `sapply()` : a version of lapply. Can simplify S the result so it's not presented as a list.
- vapply()` : has a pre-specified type of return value(V).
- *replicate* : run a function several times. Usually used with generation of random variables.
- `mapply()`: multivariate version of sapply. Arguments can be recycled.
- `rapply()` : recursive version of lapply.

## Using `apply()`

```{r, results = FALSE}
apply(Chicago_F, 1, mean)
```
check:

```{r}
mean(Chicago_F["DaysWithPrecip", ])
```
Analyze one city:

```{r, results = FALSE}
apply(Chicago_F, 1, max)
apply(Chicago_F, 1, min)
```
for practice:

```{r, results = FALSE}
apply(Chicago_F, 2, max) #doesn't make sense, but good practice
apply(Chicago_F, 2, min)
```
Comparison:

```{r, results = FALSE}
apply(Chicago_F, 1, mean)
apply(Houston_F, 1, mean)
apply(NewYork_F, 1, mean)
apply(SanFrancisco_F, 1, mean)
```
## Recreating the apply function with loops
Find the mean of each row : #1- via loop. 
```{r, results = FALSE}
output = NULL
for(i in 1:5) { #run cycle
   output[i] = mean(Chicago_F[i,])
}

names(output) = rownames(Chicago_F)
```
2: via apply function:

```{r , results = FALSE}
apply(Chicago_F, 1, mean)
```

## Using `lapply()`

```{r, results = FALSE}
t(Chicago_F)
weather_F
t(weather_F$Chicago)
newlist = lapply(weather_F, t) #list (t(weather_F$Chicago), t(weather_F$NewYork)...)
```
Example 2:

```{r , results = FALSE}
rbind(Chicago_F, NewRow = 1:12)
lapply(weather_F, rbind, NewRow = 1:12)
```
Example 3:

```{r, results = FALSE}
rowMeans(Chicago_F) # = apply(Chicago_F, 1, mean)
lapply(weather_F, rowMeans)
```
- `rowMeans()`
- `colMeans()`
- `rowSums()`
- `colSums()`

## Combining `lapply()` with the `[ ]`operator:

```{r, results = FALSE}
weather_F[[1]][1, 1]
weather_F$Chicago[1, 1]
lapply(weather_F, "[", 1, 1)
lapply(weather_F, "[", 1, )
lapply(weather_F, "[", , 3)
```

