---
title: "R_advanced_P3"
subtitle : "Play around with Factor"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import Data

```{r}
fin = read_csv("./data/Future500.csv")
head(fin)
tail(fin)
str(fin)
summary(fin)
```
Changing  from non-factor to factor:

```{r}
fin$ID = factor(fin$ID)
#summary(fin)
fin$Inception = factor(fin$Inception)
#str(fin)
```

## Factor Variable Trap (FVT)

### *Converting into Numerics For Chararcters:*

```{r}
a = c("12", "13", "14", "12", "12")
#typeof(a)
b = as.numeric(a)
#typeof(b)
```

### *Converting into Numerics For Factors:*

```{r}
z = factor(c("12", "13", "14", "12", "12"))
typeof(z)
y = as.numeric(z) #Wrong result
```
Correct way :
```{r}
x = as.numeric(as.character(z))
typeof(x)
```

## `sub()` and `gsub()`

```{r}
#?sub()
fin$Expenses = gsub(" Dollars", "", fin$Expenses)
fin$Expenses = gsub(",", "", fin$Expenses)
head(fin)
str(fin) #expenses no longer a factor but character variable

```
When including a special symbol : "$"

```{r}
fin$Revenue = gsub("\\$", "", fin$Revenue)
fin$Revenue = gsub(",", "", fin$Revenue)
head(fin)
fin$Growth = gsub("%", "", fin$Growth)
```

Convert those three variables into numeric:

```{r}
fin$Expenses = as.numeric(fin$Expenses)
fin$Revenue = as.numeric(fin$Revenue)
fin$Growth = as.numeric(fin$Growth)
str(fin)
summary(fin)
```

## Dealing with Missing Data:

1. Predict with 100% accuracy
2. Leave as it is
3. Remove from record entirely
4. Replace with mean or median(popular)
5. Fill in by exploring correlations and similarities
6. Introduce dummy variable for "Missingness"


### What is an NA? - "Not Available/Missing Value"

TRUE = 1
FALSE = 0
NA

```{r}
head(fin, 24)
fin[!complete.cases(fin), ] 
```

Update Import:

```{r}
fin1 = read.csv("./data/Future500.csv", na.strings = c(""))
head(fin1, 24)
fin1[!complete.cases(fin1), ] 
```

### Filtering: Non-missing Data with `which()`

```{r}
head(fin)
fin[fin$Revenue == 9746272, ] 
#will compare NA values with 9746272, so produced mysterious NAs.
which(fin$Revenue == 9746272)
#?which()
fin[which(fin$Revenue == 9746272), ]
#^ select the exact row 
fin[fin$Employees == 45, ]
fin[which(fin$Employees == 45), ]
```

