---
title: "R_advanced_P5_Apply"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Import Data

```{r}
Chicago_F = read.csv("./data/Weather Data/Chicago-F.csv", row.names = 1)
NewYork_F = read.csv("./data/Weather Data/NewYork-F.csv", row.names = 1)
Houston_F = read.csv("./data/Weather Data/Houston-F.csv", row.names = 1)
SanFrancisco_F = read.csv("./data/Weather Data/SanFrancisco-F.csv", row.names = 1)
```

```{r}
is.data.frame(Chicago_F)
```
### Let's convert to matrices:

```{r}
Chicago_F = as.matrix(Chicago_F)
NewYork_F = as.matrix(NewYork_F)
Houston_F = as.matrix(Houston_F)
SanFrancisco_F = as.matrix(SanFrancisco_F)
```

check:

```{r}
is.matrix(Chicago_F)
```

## Put all of these data into a list:

```{r}
weather_F = list(Chicago = Chicago_F, Houston = Houston_F, NewYork = NewYork_F, SanFrancisco = SanFrancisco_F) 
```

Try

```{r}
weather_F[[3]] %>% knitr::kable()
weather_F$NewYork %>% knitr::kable()
```

## Apply Family
- *apply* : use on a matrix, either the rows or columns.
- *tapply* : use on a vector to extract subgroups and apply a function to them by - use on data frams. Same concept as in group by in SQL.
- *eapply* : use on an environment.
- *lapply* : apply a function to elements of a list.
- *sapply* : a version of lapply. Can simplify S the result so it's not presented as a list.
- *vapply* : has a pre-specified type of return value(V).
- *replicate* : run a function several times. Usually used with generation of random variables.
- *mapply*: multivariate version of sapply. Arguments can be recycled.
- *rapply* : recursive version of lapply.